---
title: "LaneMiles_models"
output: html_document
date: "2026-06-05"
---

##Section 1: Library Loading
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("lfe")
library("dplyr")
library("ggplot2")
library("tidyr")
library("tidycensus")
library("sf")
library("tidyverse")
library("tibble")
library("lubridate")
library("psych")
library("broom")
library("lmtest")
library("plm")
library("tinytex")
library("modelsummary")
library("scales")
library("stargazer")
library("foreign")
library("readxl")
library("units")
library("ipumsr")
library("plm")

```

##Section 2: Read Databases
This section reads and arranges the data that we clean and prepared from the previous R markdown file. It's reading a dataset where each row is a County by year and by type of road (IH = Interstate Highway; MRU: Major Road Urban). Data on VMT for Caltrans PRD is given in the thousand miles, so it's being multiplied by 1000 to have it in the same units as pre 2003 numbers. It's then inflating previous household income numbers to have the same base year as 2020.
```{r, results = FALSE, warning=FALSE}

#Set working directory

setwd("C:/Users/afr/OneDrive - San Diego Association of Governments/Data Science - Documents/Economics/FY2026_27 Projects/Misc Projects/Elasticity of Lane Miles")

hpms_ipums <- read.csv("2_Data Cleaning/Code_Output/hpms_ipums_counties_8323_2026.csv")

hpms_ipums[ , 6:17] <- lapply(hpms_ipums[ , 6:17], as.numeric)
hpms_ipums[ , 19:21] <- lapply(hpms_ipums[ , 19:21], as.numeric)

hpms_ipums2 <- hpms_ipums %>%
  mutate(vmt_prd = ifelse(year >=2003, vmt_prd*1000, vmt_prd), vmt_fhwa = ifelse(year >=2003 & f_system == "ih", vmt_fhwa*1000, vmt_fhwa), hh_incr = ifelse(year == 1983, median_household_income*3.0593, ifelse(year == 1993, median_household_income*2.1087,ifelse(year == 2003, median_household_income*1.6560,ifelse(year == 2013, median_household_income*1.3080, median_household_income))))) #Need to review $ year to inflate historical value


hpms_ipums2 <- hpms_ipums2 %>%
  #filter(!is.na(vmt_fhwa)) %>%
  #filter(vmt_prd > 0) %>%
  rename(emp = s_emp,
         poverty = s_poverty)


hpms_ipums$year <- as.numeric(as.character(hpms_ipums$year))
hpms_ipums2$year <- as.numeric(as.character(hpms_ipums2$year))
hpms_ipums$county_fips <- as.factor(hpms_ipums$county_fips)
hpms_ipums2$county_fips <- as.factor(hpms_ipums2$county_fips)

hpms_ih <- hpms_ipums2 %>% filter(f_system == "ih") 
hpms_mru <- hpms_ipums2 %>% filter(f_system == "mru")

hpms_ih_fe <- pdata.frame(hpms_ih, index = c("county_fips", "year"))
hpms_mru_fe <- pdata.frame(hpms_mru, index = c("county_fips", "year"))


```

## Section 3: Models with socioeconomic variables
##Model Fixed Effects (PRD no 83-93 data)
```{r, results = FALSE}
hpms_ih_fe_prd <- hpms_ih_fe %>% 
  filter(year %in% c("2003", "2013", "2023")) %>%
  filter(!is.na(vmt_prd))

hpms_mru_fe_prd <- hpms_mru_fe %>% 
  filter(year %in% c("2003", "2013", "2023"))%>%
  filter(!is.na(vmt_prd))


#step 1 : Estimate the FE model
hpms_fe_ih_prd <- plm(log(vmt_prd) ~ log(lane_miles_prd), data = hpms_ih_fe_prd[hpms_ih_fe_prd$vmt_prd>0,], model = "within", effect = "time")
hpms_fe_mru_prd <- plm(log(vmt_prd) ~ log(lane_miles_prd), data = hpms_mru_fe_prd[hpms_mru_fe_prd$vmt_prd>0,], model = "within", effect = "time")

hpms_fe_ih_prd_pop <- plm(log(vmt_prd) ~ log(lane_miles_prd)+ log(population), data = hpms_ih_fe_prd[hpms_ih_fe_prd$vmt_prd>0,], model = "within", effect = "time")
hpms_fe_mru_prd_pop <- plm(log(vmt_prd) ~ log(lane_miles_prd)+ log(population), data = hpms_mru_fe_prd[hpms_mru_fe_prd$vmt_prd>0,], model = "within", effect = "time")

hpms_fe_ih_prd_sc <- plm(log(vmt_prd) ~ log(lane_miles_prd)+ log(population)+ log(hh_incr) + poverty + educ + emp + log(drivers_licence) + log(total_veh), data = hpms_ih_fe_prd[hpms_ih_fe_prd$vmt_prd>0,], model = "within", effect = "time")
hpms_fe_mru_prd_sc <- plm(log(vmt_prd) ~ log(lane_miles_prd)+ log(population)+ log(hh_incr) + poverty + educ + emp + log(drivers_licence) + log(total_veh), data = hpms_mru_fe_prd[hpms_mru_fe_prd$vmt_prd>0,], model = "within", effect = "time")

hpms_fe_ih_prd_county <- plm(log(vmt_prd) ~ log(lane_miles_prd), data = hpms_ih_fe_prd[hpms_ih_fe_prd$vmt_prd>0,], model = "within", effect = "twoways")
hpms_fe_mru_prd_county <- plm(log(vmt_prd) ~ log(lane_miles_prd), data = hpms_mru_fe_prd[hpms_mru_fe_prd$vmt_prd>0,], model = "within", effect = "twoways")

hpms_fe_ih_prd_pop_county <- plm(log(vmt_prd) ~ log(lane_miles_prd)+ log(population), data = hpms_ih_fe_prd[hpms_ih_fe_prd$vmt_prd>0,], model = "within", effect = "twoways")
hpms_fe_mru_prd_pop_county <- plm(log(vmt_prd) ~ log(lane_miles_prd)+ log(population), data = hpms_mru_fe_prd[hpms_mru_fe_prd$vmt_prd>0,], model = "within", effect = "twoways")

hpms_fe_ih_prd_sc_county <- plm(log(vmt_prd) ~ log(lane_miles_prd)+ log(population)+ log(hh_incr) + poverty + educ + emp , data = hpms_ih_fe_prd[hpms_ih_fe_prd$vmt_prd>0,], model = "within", effect = "twoways")
hpms_fe_mru_prd_sc_county <- plm(log(vmt_prd) ~ log(lane_miles_prd)+ log(population)+ log(hh_incr) + poverty + educ + emp , data = hpms_mru_fe_prd[hpms_mru_fe_prd$vmt_prd>0,], model = "within", effect = "twoways")


#Create the list for model comparison table
IH_models <- list(
  "(1)" = hpms_fe_ih_prd,
  "(2)" = hpms_fe_ih_prd_pop,
  "(3)" = hpms_fe_ih_prd_sc,
  "(4)" = hpms_fe_ih_prd_county,
  "(5)" = hpms_fe_ih_prd_pop_county,
  "(6)" = hpms_fe_ih_prd_sc_county)

MRU_models <- list(
  "(1)" = hpms_fe_mru_prd,
  "(2)" = hpms_fe_mru_prd_pop,
  "(3)" = hpms_fe_mru_prd_sc,
  "(4)" = hpms_fe_mru_prd_county,
  "(5)" = hpms_fe_mru_prd_pop_county,
  "(6)" = hpms_fe_mru_prd_sc_county)


# 4. Generate the Table with Driscoll-Kraay Robust SEs
rows <- tribble(
  ~term,                            ~`(1)`, ~`(2)`, ~`(3)`, ~`(4)`, ~`(5)`, ~`(6)`,
  "Time Fixed Effects",             "Y",    "Y",    "Y",    "Y",    "Y",    "Y",
  "County Fixed Effects",           "",     "",     "",     "Y",    "Y",    "Y",
  "Socioeconomic Characteristics",  "",     "",     "Y",    "",     "",     "Y"
)

modelsummary(IH_models, 
             stars = TRUE, 
             coef_map = c("log(lane_miles_prd)" = "log(Lane Miles)",
                          "log(population)"     = "log(Population)"),
             add_rows = rows,
             gof_omit = 'AIC|BIC|Log|RMSE',
             title = "VMT AS A FUNCTION OF INTERSTATE HIGHWAY LANE MILES",
             notes = "Note: Control variables for socioeconomic characteristics include: income, employment rate, poverty rate, and educational employment. Data source: Caltrans PRD")

modelsummary(MRU_models, 
             stars = TRUE, 
             coef_map = c("log(lane_miles_prd)" = "log(Lane Miles)",
                          "log(population)"     = "log(Population)"),
             add_rows = rows,
             gof_omit = 'AIC|BIC|Log|RMSE',
             title = "VMT AS A FUNCTION OF MAJOR ROAD URBAN LANE MILES",
             notes = "Note: Control variables for socioeconomic characteristics include: income, employment rate, poverty rate, and educational employment. Data source: Caltrans PRD")


```

#Predictions for PRD models
```{r, results = FALSE}
#Predict regressions with only time fixed effects for IH models
hpms_ih_fe_prd_panel <- pdata.frame(
  hpms_ih_fe_prd %>%
    select(-urban_pop_share, -median_household_income, -ct_yr2, -ua_pop) %>%
    filter(!is.na(vmt_prd), vmt_prd > 0),
  index = c("county_fips", "year")
)

pred_vals1 <- predict(hpms_fe_ih_prd, newdata = hpms_ih_fe_prd_panel)
pred_vals2 <- predict(hpms_fe_ih_prd_pop, newdata = hpms_ih_fe_prd_panel)  
pred_vals3 <- predict(hpms_fe_ih_prd_sc, newdata = hpms_ih_fe_prd_panel)
pred_vals4 <- predict(hpms_fe_ih_prd_county, newdata = hpms_ih_fe_prd_panel)
pred_vals5 <- predict(hpms_fe_ih_prd_pop_county, newdata = hpms_ih_fe_prd_panel)  
pred_vals6 <- predict(hpms_fe_ih_prd_sc_county, newdata = hpms_ih_fe_prd_panel)

# Attach predictions to your dataframe
hpms_ih_fe_prd_panel$hpms_fe_prd <- exp(as.numeric(pred_vals1))
hpms_ih_fe_prd_panel$hpms_fe_prd_pop <- exp(as.numeric(pred_vals2))
hpms_ih_fe_prd_panel$hpms_fe_prd_sc <- exp(as.numeric(pred_vals3))
hpms_ih_fe_prd_panel$hpms_fe_prd_county <- exp(as.numeric(pred_vals4))
hpms_ih_fe_prd_panel$hpms_fe_prd_pop_county <- exp(as.numeric(pred_vals5))
hpms_ih_fe_prd_panel$hpms_fe_prd_sc_county <- exp(as.numeric(pred_vals6))


#Predict regressions with only time fixed effects for MRU models
hpms_mru_fe_prd_panel <- hpms_mru_fe_prd %>%
  select(-urban_pop_share, -median_household_income, -ct_yr2, -ua_pop) %>%
  filter(!is.na(vmt_prd)) %>%
  filter(vmt_prd > 0)

# Convert data to pdata.frame using SAME index used in model estimation
new_pdf_mru <- pdata.frame(hpms_mru_fe_prd_panel, index = c("county_fips", "year"))

pred_vals1 <- predict(hpms_fe_mru_prd, newdata = new_pdf_mru)
pred_vals2 <- predict(hpms_fe_mru_prd_pop, newdata = new_pdf_mru)  
pred_vals3 <- predict(hpms_fe_mru_prd_sc, newdata = new_pdf_mru)
pred_vals4 <- predict(hpms_fe_mru_prd_county, newdata = new_pdf_mru)
pred_vals5 <- predict(hpms_fe_mru_prd_pop_county, newdata = new_pdf_mru)  
pred_vals6 <- predict(hpms_fe_mru_prd_sc_county, newdata = new_pdf_mru)

# Attach predictions to your data frame
hpms_mru_fe_prd_panel$hpms_fe_prd <- exp(as.numeric(pred_vals1))
hpms_mru_fe_prd_panel$hpms_fe_prd_pop <- exp(as.numeric(pred_vals2))
hpms_mru_fe_prd_panel$hpms_fe_prd_sc <- exp(as.numeric(pred_vals3))
hpms_mru_fe_prd_panel$hpms_fe_prd_county <- exp(as.numeric(pred_vals4))
hpms_mru_fe_prd_panel$hpms_fe_prd_pop_county <- exp(as.numeric(pred_vals5))
hpms_mru_fe_prd_panel$hpms_fe_prd_sc_county <- exp(as.numeric(pred_vals6))


#Estimate Residuals
hpms_ih_fe_prd_panel$resid_ih_1 <- residuals(hpms_fe_ih_prd)
hpms_ih_fe_prd_panel$resid_ih_2 <- residuals(hpms_fe_ih_prd_pop)
hpms_ih_fe_prd_panel$resid_ih_3 <- residuals(hpms_fe_ih_prd_sc)
hpms_ih_fe_prd_panel$resid_ih_4 <- residuals(hpms_fe_ih_prd_county)
hpms_ih_fe_prd_panel$resid_ih_5 <- residuals(hpms_fe_ih_prd_pop_county)
hpms_ih_fe_prd_panel$resid_ih_6 <- residuals(hpms_fe_ih_prd_sc_county)


hpms_mru_fe_prd_panel$resid_ih_1 <- residuals(hpms_fe_mru_prd)
hpms_mru_fe_prd_panel$resid_ih_2 <- residuals(hpms_fe_mru_prd_pop)
hpms_mru_fe_prd_panel$resid_ih_3 <- residuals(hpms_fe_mru_prd_sc)
hpms_mru_fe_prd_panel$resid_ih_4 <- residuals(hpms_fe_mru_prd_county)
hpms_mru_fe_prd_panel$resid_ih_5 <- residuals(hpms_fe_mru_prd_pop_county)
hpms_mru_fe_prd_panel$resid_ih_6 <- residuals(hpms_fe_mru_prd_sc_county)

hpms_prd_fe_panel <- rbind(hpms_ih_fe_prd_panel, hpms_mru_fe_prd_panel)

setwd("C:/Users/afr/OneDrive - San Diego Association of Governments/Data Science - Documents/Economics/FY2026_27 Projects/Misc Projects/Elasticity of Lane Miles")
path_out <- "4_Outputs"
write.csv(hpms_prd_fe_panel, file.path(path_out, "hpms_prd_fe_pred.csv"), row.names=FALSE)

```

## Section 3: Models with socioeconomic variables
##Model Fixed Effects (FHWA)
```{r, results = FALSE}
hpms_ih_fe_fhwa <- pdata.frame(hpms_ih_fe_prd, index = c("county_fips", "year"))
hpms_mru_fe_fhwa <- pdata.frame(hpms_mru_fe_prd, index = c("county_fips", "year"))

hpms_fe_ih_fhwa <- plm(log(vmt_fhwa) ~ log(lane_miles_fhwa), data = hpms_ih_fe_fhwa[hpms_ih_fe_fhwa$vmt_fhwa>0,], model = "within", effect = "time")
hpms_fe_mru_fhwa <- plm(log(vmt_fhwa) ~ log(lane_miles_fhwa), data = hpms_mru_fe_fhwa[hpms_mru_fe_fhwa$vmt_fhwa>0,], model = "within", effect = "time")

hpms_fe_ih_fhwa_pop <- plm(log(vmt_fhwa) ~ log(lane_miles_fhwa)+ log(population), data = hpms_ih_fe_fhwa[hpms_ih_fe_fhwa$vmt_fhwa>0,], model = "within", effect = "time")
hpms_fe_mru_fhwa_pop <- plm(log(vmt_fhwa) ~ log(lane_miles_fhwa)+ log(population), data = hpms_mru_fe_fhwa[hpms_mru_fe_fhwa$vmt_fhwa>0,], model = "within", effect = "time")

hpms_fe_ih_fhwa_sc <- plm(log(vmt_fhwa) ~ log(lane_miles_fhwa)+ log(population)+ log(hh_incr) + poverty + educ + emp + log(drivers_licence) + log(total_veh), data = hpms_ih_fe_fhwa[hpms_ih_fe_fhwa$vmt_fhwa>0,], model = "within", effect = "time")
hpms_fe_mru_fhwa_sc <- plm(log(vmt_fhwa) ~ log(lane_miles_fhwa)+ log(population)+ log(hh_incr) + poverty + educ + emp + log(drivers_licence) + log(total_veh), data = hpms_mru_fe_fhwa[hpms_mru_fe_fhwa$vmt_fhwa>0,], model = "within", effect = "time")

hpms_fe_ih_fhwa_county <- plm(log(vmt_fhwa) ~ log(lane_miles_fhwa), data = hpms_ih_fe_fhwa[hpms_ih_fe_fhwa$vmt_fhwa>0,], model = "within", effect = "twoways")
hpms_fe_mru_fhwa_county <- plm(log(vmt_fhwa) ~ log(lane_miles_fhwa), data = hpms_mru_fe_fhwa[hpms_mru_fe_fhwa$vmt_fhwa>0,], model = "within", effect = "twoways")

hpms_fe_ih_fhwa_pop_county <- plm(log(vmt_fhwa) ~ log(lane_miles_fhwa)+ log(population), data = hpms_ih_fe_fhwa[hpms_ih_fe_fhwa$vmt_fhwa>0,], model = "within", effect = "twoways")
hpms_fe_mru_fhwa_pop_county <- plm(log(vmt_fhwa) ~ log(lane_miles_fhwa)+ log(population), data = hpms_mru_fe_fhwa[hpms_mru_fe_fhwa$vmt_fhwa>0,], model = "within", effect = "twoways")

hpms_fe_ih_fhwa_sc_county <- plm(log(vmt_fhwa) ~ log(lane_miles_fhwa)+ log(population)+ log(hh_incr) + poverty + educ + emp , data = hpms_ih_fe_fhwa[hpms_ih_fe_fhwa$vmt_fhwa>0,], model = "within", effect = "twoways")
hpms_fe_mru_fhwa_sc_county <- plm(log(vmt_fhwa) ~ log(lane_miles_fhwa)+ log(population)+ log(hh_incr) + poverty + educ + emp , data = hpms_mru_fe_fhwa[hpms_mru_fe_fhwa$vmt_fhwa>0,], model = "within", effect = "twoways")


#Create the list for model comparison table
IH_models <- list(
  "(1)" = hpms_fe_ih_fhwa,
  "(2)" = hpms_fe_ih_fhwa_pop,
  "(3)" = hpms_fe_ih_fhwa_sc,
  "(4)" = hpms_fe_ih_fhwa_county,
  "(5)" = hpms_fe_ih_fhwa_pop_county,
  "(6)" = hpms_fe_ih_fhwa_sc_county)

MRU_models <- list(
  "(1)" = hpms_fe_mru_fhwa,
  "(2)" = hpms_fe_mru_fhwa_pop,
  "(3)" = hpms_fe_mru_fhwa_sc,
  "(4)" = hpms_fe_mru_fhwa_county,
  "(5)" = hpms_fe_mru_fhwa_pop_county,
  "(6)" = hpms_fe_mru_fhwa_sc_county)


# 4. Generate the Table with Driscoll-Kraay Robust SEs
rows <- tribble(
  ~term,                            ~`(1)`, ~`(2)`, ~`(3)`, ~`(4)`, ~`(5)`, ~`(6)`,
  "Time Fixed Effects",             "Y",    "Y",    "Y",    "Y",    "Y",    "Y",
  "County Fixed Effects",           "",     "",     "",     "Y",    "Y",    "Y",
  "Socioeconomic Characteristics",  "",     "",     "Y",    "",     "",     "Y"
)

modelsummary(IH_models, 
             stars = TRUE, 
             coef_map = c("log(lane_miles_fhwa)" = "log(Lane Miles)",
                          "log(population)"     = "log(Population)"),
             add_rows = rows,
             gof_omit = 'AIC|BIC|Log|RMSE',
             title = "VMT AS A FUNCTION OF INTERSTATE HIGHWAY LANE MILES",
             notes = "Note: Control variables for socioeconomic characteristics include: income, employment rate, poverty rate, and educational employment. Data source: Caltrans fhwa")

modelsummary(MRU_models, 
             stars = TRUE, 
             coef_map = c("log(lane_miles_fhwa)" = "log(Lane Miles)",
                          "log(population)"     = "log(Population)"),
             add_rows = rows,
             gof_omit = 'AIC|BIC|Log|RMSE',
             title = "VMT AS A FUNCTION OF MAJOR ROAD URBAN LANE MILES",
             notes = "Note: Control variables for socioeconomic characteristics include: income, employment rate, poverty rate, and educational employment. Data source: Caltrans fhwa")

```

#Predictions for PRD models
```{r, results = FALSE}
#Predict regressions with only time fixed effects for IH models
hpms_ih_fe_fhwa_panel <- pdata.frame(
  hpms_ih_fe_fhwa %>%
    select(-urban_pop_share, -median_household_income, -ct_yr2, -ua_pop) %>%
    filter(!is.na(vmt_fhwa), vmt_fhwa > 0),
  index = c("county_fips", "year")
)

pred_vals1 <- predict(hpms_fe_ih_fhwa, newdata = hpms_ih_fe_fhwa_panel)
pred_vals2 <- predict(hpms_fe_ih_fhwa_pop, newdata = hpms_ih_fe_fhwa_panel)  
pred_vals3 <- predict(hpms_fe_ih_fhwa_sc, newdata = hpms_ih_fe_fhwa_panel)
pred_vals4 <- predict(hpms_fe_ih_fhwa_county, newdata = hpms_ih_fe_fhwa_panel)
pred_vals5 <- predict(hpms_fe_ih_fhwa_pop_county, newdata = hpms_ih_fe_fhwa_panel)  
pred_vals6 <- predict(hpms_fe_ih_fhwa_sc_county, newdata = hpms_ih_fe_fhwa_panel)

# Attach predictions to your dataframe
hpms_ih_fe_fhwa_panel$hpms_fe_fhwa <- exp(as.numeric(pred_vals1))
hpms_ih_fe_fhwa_panel$hpms_fe_fhwa_pop <- exp(as.numeric(pred_vals2))
hpms_ih_fe_fhwa_panel$hpms_fe_fhwa_sc <- exp(as.numeric(pred_vals3))
hpms_ih_fe_fhwa_panel$hpms_fe_fhwa_county <- exp(as.numeric(pred_vals4))
hpms_ih_fe_fhwa_panel$hpms_fe_fhwa_pop_county <- exp(as.numeric(pred_vals5))
hpms_ih_fe_fhwa_panel$hpms_fe_fhwa_sc_county <- exp(as.numeric(pred_vals6))


#Predict regressions with only time fixed effects for MRU models
hpms_mru_fe_fhwa_panel <- hpms_mru_fe_fhwa %>%
  select(-urban_pop_share, -median_household_income, -ct_yr2, -ua_pop) %>%
  filter(!is.na(vmt_fhwa)) %>%
  filter(vmt_fhwa > 0)

# Convert data to pdata.frame using SAME index used in model estimation
new_pdf_mru <- pdata.frame(hpms_mru_fe_fhwa_panel, index = c("county_fips", "year"))

pred_vals1 <- predict(hpms_fe_mru_fhwa, newdata = new_pdf_mru)
pred_vals2 <- predict(hpms_fe_mru_fhwa_pop, newdata = new_pdf_mru)  
pred_vals3 <- predict(hpms_fe_mru_fhwa_sc, newdata = new_pdf_mru)
pred_vals4 <- predict(hpms_fe_mru_fhwa_county, newdata = new_pdf_mru)
pred_vals5 <- predict(hpms_fe_mru_fhwa_pop_county, newdata = new_pdf_mru)  
pred_vals6 <- predict(hpms_fe_mru_fhwa_sc_county, newdata = new_pdf_mru)

# Attach predictions to your data frame
hpms_mru_fe_fhwa_panel$hpms_fe_fhwa <- exp(as.numeric(pred_vals1))
hpms_mru_fe_fhwa_panel$hpms_fe_fhwa_pop <- exp(as.numeric(pred_vals2))
hpms_mru_fe_fhwa_panel$hpms_fe_fhwa_sc <- exp(as.numeric(pred_vals3))
hpms_mru_fe_fhwa_panel$hpms_fe_fhwa_county <- exp(as.numeric(pred_vals4))
hpms_mru_fe_fhwa_panel$hpms_fe_fhwa_pop_county <- exp(as.numeric(pred_vals5))
hpms_mru_fe_fhwa_panel$hpms_fe_fhwa_sc_county <- exp(as.numeric(pred_vals6))


#Estimate Residuals
hpms_ih_fe_fhwa_panel$resid_ih_1 <- residuals(hpms_fe_ih_fhwa)
hpms_ih_fe_fhwa_panel$resid_ih_2 <- residuals(hpms_fe_ih_fhwa_pop)
hpms_ih_fe_fhwa_panel$resid_ih_3 <- residuals(hpms_fe_ih_fhwa_sc)
hpms_ih_fe_fhwa_panel$resid_ih_4 <- residuals(hpms_fe_ih_fhwa_county)
hpms_ih_fe_fhwa_panel$resid_ih_5 <- residuals(hpms_fe_ih_fhwa_pop_county)
hpms_ih_fe_fhwa_panel$resid_ih_6 <- residuals(hpms_fe_ih_fhwa_sc_county)


hpms_mru_fe_fhwa_panel$resid_ih_1 <- residuals(hpms_fe_mru_fhwa)
hpms_mru_fe_fhwa_panel$resid_ih_2 <- residuals(hpms_fe_mru_fhwa_pop)
hpms_mru_fe_fhwa_panel$resid_ih_3 <- residuals(hpms_fe_mru_fhwa_sc)
hpms_mru_fe_fhwa_panel$resid_ih_4 <- residuals(hpms_fe_mru_fhwa_county)
hpms_mru_fe_fhwa_panel$resid_ih_5 <- residuals(hpms_fe_mru_fhwa_pop_county)
hpms_mru_fe_fhwa_panel$resid_ih_6 <- residuals(hpms_fe_mru_fhwa_sc_county)

hpms_fhwa_fe_panel <- rbind(hpms_ih_fe_fhwa_panel, hpms_mru_fe_fhwa_panel)


setwd("C:/Users/afr/OneDrive - San Diego Association of Governments/Data Science - Documents/Economics/FY2026_27 Projects/Misc Projects/Elasticity of Lane Miles")
path_out <- "4_Outputs"
write.csv(hpms_fhwa_fe_panel, file.path(path_out, "hpms_fhwa_fe_pred.csv"), row.names=FALSE)

```



